---
title: "Clinical data of COVID-19 infected patients"
format: 
  revealjs:
    transition: none # Global default
editor: visual
subtite: "A case of study for critically ill patients"
css: styles.css
---

# Introduction {background-image="background.avif"}

(igual aqui habria que poner los nombres o algo supongo)

## Introduction

### Covid19

## Introduction

### Mortality

## Introduction

### Paxlovid and drugs

```{r, echo=FALSE, results='asis'}
library (grid)

img <- png::readPNG("../R/prueba.png")
grid::grid.raster(img)
```

esta foto está aqui solo para que tengamos el codigo de como poner fotos

# Dataset

## Dataset

Quick overview fail, no se como exportar las table 1 para ponerlos aqui

```{r, echo=FALSE, results='asis'}
#library (grid)

#img <- png::readPNG("../results/table1_describe_01.png")
#grid::grid.raster(img)
```

```{r, echo=FALSE, results='asis'}
library(tidyverse)
library(knitr)

knit("../R/04_describe.qmd")



```

# Analysis {.transition-fade}

## Analysis {.transition-fade}

::: incremental
-   Initial exploration: What factors we think might be related to the outcome?
-   Statistical exploration: What factors are statistically related to the outcome?
-   Paxlovid exploration: Can we get the same result as the main paper with this data?
:::

## Analysis - Initial exploration

::: nonincremental
-   Initial exploration: What factors we think might be related to the outcome?
:::

## Analysis - Initial exploration

::: nonincremental
-   Initial exploration: What factors we think might be related to the outcome?
    -   Age
    -   symptoms
    -   time in UCI
    -   other disease
    -   extra correlations
:::

## Analysis - Initial exploration

What factors we think might be related to the outcome?

## Analysis - Statistical exploration

What factors are statistically related to the outcome?

```{r}
library(tidyverse)
#library(knitr)

load("../results/significant_variables.RData")
significant_variables_direct_corr <- significant_variables |> 
  arrange(desc(estimate))

significant_variables_inverse_corr <- significant_variables |> 
  arrange(estimate)

#kable(significant_variables_direct_corr, caption = "Tabla de Estimaciones")
head(significant_variables_direct_corr, n=10)
#head(significant_variables_inverse_corr)
```

## Analysis - Statistical exploration

What factors are statistically related to the outcome?

```{r}
library(tidyverse)
library(knitr)

#kable(significant_variables_direct_corr, caption = "Tabla de Estimaciones")
head(significant_variables_inverse_corr, n=10)
#head(significant_variables_inverse_corr)
```

## Analysis - Statistical exploration

Variables with direct correlation

## Analysis - Statistical exploration

Variables with inverse correlation

por aqui tendria pensado poner lo del analisis quitando a la gente con sangradosss, y que nos de pie a explicar todo el tema de la trombosis y tal

## Analysis - Paxoliv exploration

results about paxlovid?
