---
title: "Clinical data of COVID-19 infected patients"
format: 
  revealjs:
    transition: none # Global default
editor: visual
subtite: "A case of study for critically ill patients"
css: styles.css
---

# Introduction {background-image=background.avif}

## Introduction

### Covid19

## Introduction

### Mortality

## Introduction

### Paxlovid and drugs

```{r, echo=FALSE, results='asis'}
library (grid)

img <- png::readPNG("../R/prueba.png")
grid::grid.raster(img)
```


# Dataset

## Dataset
Quick overview

# Our hypothesis {.transition-fade}

## Our Hypothesis {.transition-fade}

::: {.incremental} 

- What factors we think might the related to the outcome?
- What factors are statistically related to the outcome?
- Is there a difference on the use of Paxlovid?

:::

## Our Hypothesis

::: {.nonincremental}

- What factors we think might the related to the outcome?

:::

## Our Hypothesis

::: {.nonincremental}

- What factors we think might the related to the outcome?
  - age
  - symptoms
  - time in UCI
  - other disease
  - extra correlations

:::

## What factors we think might the related to the outcome?
age, simptons time in UCI, other disease

## What factors are statistically related to the outcome?

```{r}
library(tidyverse)
#library(knitr)

load("../results/significant_variables.RData")
significant_variables_direct_corr <- significant_variables |> 
  arrange(desc(estimate))

significant_variables_inverse_corr <- significant_variables |> 
  arrange(estimate)

#kable(significant_variables_direct_corr, caption = "Tabla de Estimaciones")
head(significant_variables_direct_corr, n=10)
#head(significant_variables_inverse_corr)
```

## What factors are statistically related to the outcome?

```{r}
library(tidyverse)
library(knitr)

#kable(significant_variables_direct_corr, caption = "Tabla de Estimaciones")
head(significant_variables_inverse_corr, n=10)
#head(significant_variables_inverse_corr)
```

## What factors are statistically related to the outcome?
Variables with direct correlation

## What factors are statistically related to the outcome?
Variables with inverse correlation




